Stability of variational quantum dynamics with Gaussian approximants

### 2013-07-15 paper.tex
morning writing

We need a way to discretise the function $g$.  There are two usual ways to do this: sampling on a grid, or taking components along a fixed basis.  The size of state space rules these out, so we must sample $g$ at a set of irregularly spaced amplitudes $A=\{α₁,…,α_R\}$.  Equivalently, we take components of $f$ in function space along the set of vectors $χ_r(α)=w_rδ(α-α_r)$.  Here, $w_r$ is a real weight, intended to compensate for the irregularity of the sample set, so that the sampled vector $\vec g=\bigl(w₁g(α₁),…,w_Rg(α_R)\bigr)^{\rm T}$ has a similar norm to the function $g$.

@@

The norm of $Gg$ is 1, so this suggests the singular values of $G$ are $π^{-½}$.  Numerical experiments in the script {\tt sve.m} support this.  The operator $G$ was discretised on a grid in phase space, with spacing $h=0.3$, and various extents shown in the graphs.  Ket space was discretised by taking inner products with the number states $|0〉$ through $|10〉$.  The graph below shows the singular values of the discretised operator converging to $π^{-½}$, as the extent of the grid passes through $|α|²=10$.

>>>>

We need a way to discretise the function $g$.  There are two usual ways to do this: sampling on a grid, or taking components along a fixed basis.  The size of state space rules these out, so we must sample $g$ at a set of irregularly spaced amplitudes $A=\{α₁,…,α_R\}$, corresponding to a row of kets $|A〉=\bigl(|α₁〉\ ⋯\ |α_R〉\bigr).  Equivalently, we take components of $f$ in function space along the set of vectors $χ_r(α)=w_rδ(α-α_r)$.  Here, $w_r$ is a real weight, intended to compensate for the irregularity of the sample set, so that the sampled vector $\vec g=\bigl(w₁g(α₁),…,w_Rg(α_R)\bigr)^{\rm T}$ has a similar norm to the function $g$.

@@

The norm of $Gg$ is 1, so this suggests the singular values of $G$ are $π^{-½}$.  This can also be seen from the expansions of coherent states.  The normalised entire expansion of $|α〉$ is
$$g(β*)=〈β|α〉=e^{-½|α|²-½|β|²+β*α},$$
whence
$$\|g\|²=\int |g(β)|²\,d²β=\int e^{-|β-α|²}\,d²β=π.$$
The norm of $|α〉$ is of course 1.  If we used the Bargmann space, we could easily absorb the $π$ into the inner product, and make $G$ a generalised projector.

The lemma gives a very simple singular value expansion of $G$.  Because there is only one nonzero singular value, any orthonormal basis of ket space will do as the left singular vectors.  Let's choose the number states for familiarity.  Then
$$Gg=∑_n |n〉π\int ν_n*(α)g(α)\,d²α=|N〉π\int ν*(α)g(α),$$
where $|N〉=\bigl(|0〉\ |1〉\ |2〉\ …\bigr)$ and $ν(α)=\bigl(ν₀(α)\ ν₁(α)\ …\bigr)$ could be thought of as infinite matrices whose columns are the left singular vectors $|n〉$ and their normalised entire expansions, the right singular vectors $v_n$.  The pesudoinverse, 
$$G⁺|ψ〉=νπ〈N|ψ〉= ∑_n ν_nπ〈n|ψ〉,$$
maps $|ψ〉$ to its normalised entire expansion.  Expanding over number states is useful because the maximum number in a system is usually limited by conservation of energy.  In real systems, the components decay exponentially with $n$, and become smaller than numerical rounding errors at reasonably small numbers.

Numerical experiments in the script {\tt sve.m} support this.  The operator $G$ was discretised on a grid in phase space, with spacing $h=0.3$, and various extents shown in the graphs.  Ket space was discretised by taking inner products with the number states $|0〉$ through $|10〉$.  The graph below shows the singular values of the discretised operator converging to $π^{-½}$, as the extent of the grid passes through $|α|²=10$.
